From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-06-05 dictionary.tex
Minor improvements.

The process of creating a category from the group of quasicharacters of a $p$-adic  torus informs our choice of the term quasicharacter sheaf in this paper.
We also want to situate our terminology in a historical context.
It is not uncommon to refer to local systems satisfying \eqref{introbox} on a connected, commutative algebraic group as character sheaves;
see for example, \cite{kamgarpour:09a}*{Intro}.
As we have indicated, quasicharacter sheaves evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
However, the main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on reductive groups
over algebraically closed fields with finite cyclic component groups in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, such groups are extensions of finite cyclic groups by algebraic tori,
in which case it is not difficult to relate Frobenius-stable character sheaves to our character sheaves.
%The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0({\bar G}),\EEx)^{\Weil{}}$ do not arise in that context.

@@

\begin{proof}
Let $\cs{L}$ be a character sheaf on $G$.
Using Proposition~\ref{prop:SGiso}, let $[\alpha \oplus \beta \oplus 0]\in \Hh^2(E^\bullet_G)$ be the corresponding class and let  $a : \bG \times \bG \to \EEx$ and $b : \bG \to \EEx$ be multiplicative versions of $\alpha$ and $\beta$, following the notation at the beginning of this section.
Then $\trFrob{\cs{L}}(x) = b(x)$ for $x\in G(\Fq)$.
The identity component $G^0$ of $G$ may be identified with $\Spec{\Fq}$, 
By \cite{beilinson-bernstein-deligne:81a}*{Rem. 5.1.2},
$\cs{L}$ descends to $G$ if and only if $b(x) \in \EEx$ is bounded, for all $x$.

@@

Proposition~\ref{prop:bounded} shows that the full subcategory
$\bCS(G) \subset \CS(G)$ is not an equivalence, for smooth commutative group schemes $G$.
This simple observation reveals the necessity of working with Weil sheaves in Definition~\ref{def:CS}: one cannot geometrize all characters of $G(\Fq)$ using local systems on $G$, for general smooth commutative groups schemes $G$. 
For example, consider the case when $G$ is the discrete \'etale group scheme $\ZZ$.
If $\chi : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi(1) = \ell$
and if $\cs{L}$ is a character sheaf on $G$ in the isomorphism class
corresponding to $\chi$ under Theorem~\ref{thm:main},
then $\cs{L}$ does not descend to $G$ since the image of $\chi$ is not
bounded.

>>>>

The process of creating a category from the group of quasicharacters of a $p$-adic  torus informs our choice of the term quasicharacter sheaf in this paper.
We also want to situate our terminology in a historical context.
It is not uncommon to refer to local systems satisfying \eqref{introbox} on a connected, commutative algebraic group as character sheaves;
see for example, \cite{kamgarpour:09a}*{Intro}.
As we have indicated, quasicharacter sheaves evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
However, the main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on reductive groups
over algebraically closed fields with finite cyclic component groups in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, such groups are extensions of finite cyclic groups by algebraic tori,
in which case it is not difficult to relate Frobenius-stable character sheaves to our character sheaves.
The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0({\bar G}),\EEx)^{\Weil{}}$ do not arise in that context.

@@

\begin{proof}
Let $\cs{L}$ be a character sheaf on $G$.
Using Proposition~\ref{prop:SGiso}, let $[\alpha \oplus \beta \oplus 0]\in \Hh^2(E^\bullet_G)$ be the corresponding class and let  $a : \bG \times \bG \to \EEx$ and $b : \bG \to \EEx$ be multiplicative versions of $\alpha$ and $\beta$, following the notation at the beginning of this section.
Then $\trFrob{\cs{L}}(x) = b(x)$ for $x\in G(\Fq)$.
To prove the corollary we will show that $\cs{L}$ descends to $G$ if and only if $b(x)\in \mathbb{\bar Z}_{\ell}^\times$ for every $x\in G(\Fq)$.

The identity component $G^0$ of $G$ may be identified with $\Spec{\Fq}$, 
By \cite{beilinson-bernstein-deligne:81a}*{Rem. 5.1.2},
$\cs{L}$ descends to $G$ if and only if $b(x) \in \EEx$ is bounded, for all $x$.

@@

Proposition~\ref{prop:bounded} shows that the full subcategory
$\bCS(G) \subset \CS(G)$ is not an equivalence, for smooth commutative group schemes $G$.
This simple observation reveals the necessity of working with Weil sheaves in Definition~\ref{def:CS}: one cannot geometrize all characters of $G(\Fq)$ using local systems on $G$, for general smooth commutative groups schemes $G$.

For example, consider the case when $G$ is the discrete \'etale group scheme $\ZZ$.
If $\chi : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi(1) = \ell$
and if $\cs{L}$ is a character sheaf on $G$ in the isomorphism class
corresponding to $\chi$ under Theorem~\ref{thm:snake},
then $\cs{L}$ does not descend to $G$ since the image of $\chi$ is not
bounded.
